---
title: "Gentrification other"
author: "Michael Mann"
date: "January 4, 2018"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r  comment = NA, echo = FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(rgdal)
library(sp)
library(RColorBrewer)
library(tmap)
#library(corrplot)
#library(normtest)
#library(psych)
#library(leaps)
#library(lmtest)
library(car)
#library(sphet)
#library(spgwr)
#library(spData)
#library(classInt)
#library(spdep)
 
```

```{r  comment = NA, message=FALSE, warning=FALSE}
setwd("/Users/mmann/Desktop/DC_gentrification/")
```
### Data prep
```{r comment = NA, message=FALSE, warning=FALSE}
gen<-readOGR("./gen_raw_data.GeoJSON")
names(gen)
gen$change_vcr<-as.numeric(as.character((gen$change_vcr))) #  change data type 
gen$change_ren<-as.numeric(as.character((gen$change_ren)))
gen$change_hhs<-as.numeric(as.character((gen$change_hhs)))
##### First Order Queen Contiguity matrix

proj4string(gen)<-CRS("+proj=longlat +datum=NAD83 +no_defs +ellps=GRS80 +towgs84=0,0,0") # Create projection  in case .pjr file doesn't come alone with the shp file.

#gen_NAD<-spTransform(gen, CRS("+init=ESRI:102685")) # Reproject to ESRI:102685 (NAD 1983 Stateplane Maryland FIPS 1900)

coords<-coordinates(gen) 

```


### Regression
```{r  comment = NA, message=FALSE, warning=FALSE}
for1<-change_inc~p_chg_wh+change_ren+p_chg_pvt+p_chg_edc # first formula
```
### Basic OLS
```{r  comment = NA, message=FALSE, warning=FALSE}
model.lm<-lm(for1, data =gen) # regression using first formula
summary(model.lm)
```
******

### Change in White pop OLS Prediction  

This shows how change in % white population is estimated to effect change in income, holding all other variables at thier mean.
```{r  comment = NA, message=FALSE, warning=FALSE}
# coeff for chg_wht * gen data + all other variables at their mean
model.lm<-lm(for1, data =gen) # regression using first formula
gen$wht_chng_pred = model.lm$coefficients['(Intercept)']+model.lm$coefficients['p_chg_wh']*gen$p_chg_wh+model.lm$coefficients['change_ren']*mean(gen$change_ren,na.rm=T)+model.lm$coefficients['p_chg_pvt']*mean(gen$p_chg_pvt,na.rm=T)+model.lm$coefficients['p_chg_edc']*mean(gen$p_chg_edc,na.rm=T)
qtm(gen, fill="wht_chng_pred", fill.style = "jenks", fill.palette ="RdBu", borders ="#f1f1f1")

```

******

### Change in rent and education OLS Prediction 
This shows how change in % white population and change in rents is estimated to effect change in income, holding all other variables at thier mean.

```{r  comment = NA, message=FALSE, warning=FALSE}
model.lm<-lm(for1, data =gen) # 
gen$wht_edu_chng_pred = model.lm$coefficients['(Intercept)']+model.lm$coefficients['p_chg_wh']*gen$p_chg_wh +  model.lm$coefficients['change_ren']*mean(gen$change_ren,na.rm=T)+model.lm$coefficients['p_chg_pvt']*mean(gen$p_chg_pvt,na.rm=T)+model.lm$coefficients['p_chg_edc']*gen$p_chg_edc

qtm(gen, fill="wht_edu_chng_pred", fill.style = "jenks", fill.palette ="RdBu", borders ="#f1f1f1")

```

******

### Change in rent, education, & poverty OLS Prediction 
This shows how change in % white population and change in rents is estimated to effect change in income, holding all other variables at thier mean.
```{r  comment = NA, message=FALSE, warning=FALSE}
model.lm<-lm(for1, data =gen) # 
gen$wht_edu_chng_pred = model.lm$coefficients['(Intercept)']+model.lm$coefficients['p_chg_wh']*gen$p_chg_wh +  model.lm$coefficients['change_ren']*mean(gen$change_ren,na.rm=T)+model.lm$coefficients['p_chg_pvt']*gen$p_chg_pvt+model.lm$coefficients['p_chg_edc']*gen$p_chg_edc

qtm(gen, fill="wht_edu_chng_pred", fill.style = "jenks", fill.palette ="RdBu", borders ="#f1f1f1")

```

******

###  Z-score scale 

```{r}

z.inc<-scale(gen$change_inc, scale = T, center = T)
z.wht<-scale(gen$p_chg_wh, scale = T, center = T)
z.edc<-scale(gen$p_chg_edc, scale = T, center = T)
z.pvt<-scale(gen$p_chg_pvt, scale = T, center = T)
z.ren<-scale(gen$change_ren, scale = T, center = T)

weighted.z<-data.frame((z.inc+z.wht+z.edc+z.ren)-z.pvt) 
colnames(weighted.z) <-"wgt.z"

scaled_zscore<-data.frame(cbind(z.inc,z.wht, z.edc, z.ren, z.pvt, weighted.z))
colnames(scaled_zscore) <-c("z.inc", "z.wht", "z.edc", "z.ren", "z.pvt", "wgt.z")
gen$weighted_z_sum<-scaled_zscore$wgt.z


#Jenks MAP of weighted z-score

qtm(gen, fill="weighted_z_sum", fill.style = "jenks", fill.palette ="RdBu", borders ="#f1f1f1")

```

******

### 0 to 1 Scale 

```{r  comment = NA, message=FALSE, warning=FALSE}
# Function to scale betwen 0 and 1

rescale <- function(v, lower = 0, upper = 1) {

  L <- min(v)
  H <- max(v)
  result <- (v - L) / (H - L) * (upper - lower) + lower
  return(result)
}
s.inc<-rescale(gen$change_inc)
s.wht<-rescale(gen$p_chg_wh) #scaled to 0 and 1
s.edc<-rescale(gen$p_chg_edc)
s.pvt<-rescale(gen$p_chg_pvt)
s.ren<-rescale(gen$change_ren)

weighted.s<-data.frame((s.inc+s.wht+s.edc+s.ren)-s.pvt) # Equation for weighted z

colnames(weighted.s) <-"wgt.s"
scaled_values<-data.frame(cbind(s.inc, s.wht, s.edc, s.ren, s.pvt, weighted.s))
colnames(scaled_values) <-c("s.inc", "s.wht", "s.edc", "s.ren", "s.pvt", "wgt.s")
gen$weighted_sum<-scaled_values$wgt.s

#QUANTILE MAP

qtm(gen, fill="weighted_sum", fill.style = "jenks", fill.palette ="RdBu", borders ="#f1f1f1")

```

````{r section} 
library(sf)
st_write(st_as_sf(gen), "GentrificationModels.GeoJSON")
```

